Context Inconsistency Management \\in Pervasive Systems

### 2011-04-17 paper/paper.tex
final review paper

%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
\keywords{Pervasive Computing, Ontology Model, Context Lifecycle, Inconsistency Resolution.}

@@

In the next chapter we start with an overview of the existing solutions for handling context inconsistencies and argue why these solutions do not meet the requirements of pervasive applications. Next we present Partial Constraint Checking: an algorithm for constraint checking on huge sets of contexts that is far more efficient and effective than the existing approaches. In addition we discuss a Context Inconsistency Resolution algorithm that allows for resolving conflicting context sets without human intervention. Finally we propose a whole new approach to context inconsistencies: incorporating inconsistencies in to the context model by using fuzzy ontologies. We show that this approach makes conflict detection and resolution superfluous and has lots of potential for an efficient implementation using Partial Probability Calculation.

@@

Context management for consistency however has not been adequately studied in the existing literature. None of the studies on context-awareness discusses a way for detecting context inconsistencies for reliable pervasive computing~\cite{xu:2010:PCC, bu:2006:CCM}. Even though other disciplines as artificial intelligence and software engineering conducted relate research, it does not provide adequate support for context inconsistency detection in ubiquitous computing. In addition the strategies proposed in literature for resolving context conflicts are not suited for pervasive computing. Some are based on assumptions that may not apply to general pervasive environments. Other require human participation for conflict resolution, which is usually expensive and slow for pervasive computing~\cite{xu:2010:PCC}. Finally no research has been done on the potential of fuzzy ontologies to represent context inconsistencies in the environment instead of trying to resolve them~\cite{ko:2009:IOFO}.

@@

\section{Partial Constraint Checking}
Constraint checking techniques have been extensively studied in software engineering. Existing constraint checking algorithms focus on checking software artifacts that do not change rapidly or frequently~\cite{xu:2010:PCC}. Context-aware applications require more efficient algorithms because they use a huge set of contexts, which can change very rapidly and frequently (in the range of milliseconds). An inefficient software solution for this does not only require massive computing power, but interestingly also induces a higher inconsistency detection miss rate. Because of the computing delay conflicting contexts slip through the context buffer before they are detected by the software~\cite{xu:2010:PCC}. One example of an inefficient approach is non-incremental checking: whenever there is a change in the set of software artifacts these artifacts are each checked against the entire set of consistency constraints to find out all detectable inconsistencies. An improvement to this would be incremental checking: only a subset of all constraints are checked, namely those that are affected by the specific change in the artifact set. But the real merit of Xu et \textit{al.} was replacing the traditional entire constraint checking approach by partial constraint checking based on a consistency computation tree~\cite{xu:2010:PCC}. Their idea is that constraints can be represented as trees with nodes for logical operations and leaves for specific properties of contexts or context sets. Whenever a context is added to or deleted from a context set in time, the branch corresponding to this context can respectively be added to or deleted from the tree (see Fig.~\ref{fig:cct}). 
\begin{figure}[htb]
  \centering
  \includegraphics[width=3.5in]{cons_comp_tree}
  \caption{Consistency Computation Tree: branches can be added or deleted to represent changes in a context set. Source:~\cite{xu:2010:PCC}}
  \label{fig:cct}
\end{figure}
Because intermediate values are retained in the tree nodes after calculation they can be reused whenever the tree changes. More specifically, when a branch is added, only the values of the new branch itself and of the nodes from the branch top to the tree root need to be calculated. When a branch is deleted, only the values for the nodes from the branch top to the tree root have to be recalculated.

@@

\section{Context Inconsistency Resolution}
Now that we have an efficient method for detecting inconsistencies the task of resolving the conflicts remains. As discussed above the strategies for conflict resolution in literature are not well suited for pervasive computing because their assumptions do not apply or they require human participation~\cite{xu:2010:PCC}. What is really needed is an algorithm that in the case of a conflict between two or more contexts decides which context has the highest reliability and retains that context. Bu et \textit{al.} show us that this is possible through extending the ontology based context model with additional information about the context's status and temporal properties~\cite{bu:2006:CCM}. More specifically they propose an algorithm that for every conflict set retains the context with the highest \textit{relative frequency}: frequency of context updates relative to update interval and context age (see Fig.~\ref{fig:cir}). 
\begin{figure}[htb]
  \centering
  \includegraphics[width=3.5in]{cir}
  \caption{Context inconsistency resolution in action. Source:~\cite{bu:2006:CCM}}
  \label{fig:cir}
\end{figure}
This is based on the assumption that contexts that are most stably perceived by a system are most likely to be correct. This assumption is applicable to most of the environments in which pervasive systems run and is applied in many domains as a domain specific approach~\cite{xu:2010:PCC}.

@@

\subsection{Case Study}
Suppose we want to build a context-aware application that has a certain notion of teacher's activities to assist them in the best possible way (for instance automatically opening slides when a lecture starts). In the traditional approach we could build an ontology specifying where lectures take place and that a lecture starts when the teacher appears at the lectern. If now an unpredicted observation is made (sensor cross read or teacher momentarily leaves the room), an inconsistency occurs that has to be resolved for the application to proceed with its normal operation. In the fuzzy ontology approach these unpredictable events can be modeled so that the ontology is more robust and always yields the most apt interpretation of the environment without the need for conflict resolution. A simplified hidden Markov model for this ontology is depicted in Fig.~\ref{fig:fuzzy}.
\begin{figure}[htb]
  \centering
  \includegraphics[width=3.5in]{fuzzy}
  \caption{Fuzzy ontology specifying a teacher's activities and the probabilities of different locations in the presence of these higher-level context.}
  \label{fig:fuzzy}
\end{figure}
As the numbers in this model indicate, the most probable location of a teacher while lecturing is at the lectern (L). In the aisle (A) is less probable and at the office (O) is very unlikely. For the activity of performing other work (denoted as Working in the model), the probabilities are the other way around. This fuzzy ontology model can now be used to calculate the relative probability of each higher-level context (c.q. Lecturing and Working) based a set of low level contexts observed within a certain time window. A possible set of observations could be [$L_1$, $L_2$, $L_3$, $A_1$, $A_2$]: the teacher was observed three times at the lectern and twice in the aisle in this time window. The probability of the Lecturing activity can now be calculated by multiplying the probabilities of the observations in the presence of this higher level context\footnote{We multiply the results with a constant $c$ to obtain a valid probability value in the end}: $pLecturing = .80 * .80 * .80 * .15 * .15 * c_1 = .01152 * c_1 = 0.9998$. The probability for Working can be calculated in a similar way: $pWorking = .05 * .05 * .05 * .15 * .15 * c_1 = 0.000002813 * c_1 = 0.0002$. The probabilities indicate that the teacher is still lecturing, even though the teacher being observed in the aisle conflicts with being observed at the lectern. Another example of observations could be [$L_1$, $A_1$, $A_2$, $O_1$, $O_2$]: the teacher is observed at the lectern once, twice in the aisle and twice at the office. The probabilities for Lecturing and Working are calculated as follows: $pLecturing = .80 * .15 * .15 * .05 * .05 * c_2 = 0.000045 * c_2 = 0.059$ and $pWorking = .05 * .15 * .15 * .80 * .80 * c_2 = 0.00072 * c_2 = 0.941$. The probabilities now indicate that the teacher is doing work other than lecturing, even though the observation of the teacher at the lectern is definitely conflicting with the observations at the office.
\subsection{Results}
The case study shows that a fuzzy ontology allows us to always define the most probable higher-level context given the observations, regardless of any inconsistencies in the observed context set and without the need for conflict detection and resolution.

Another observation we can do is that the hidden Markov model for a fuzzy ontology has a tree-like structure. This structure allows for very efficient incremental Partial Probability Calculation (PPC). The PCC principle works as follows: the probability of a high-level context is calculated by multiplying the probabilities of its children contexts. When a new context is added to or removed from the observation set and thus from the tree, only the probabilities of the observation's ancestors in the tree up to the root have to be recalculated. In case of adding a context, the probability of an ancestor in the tree can be recalculated by simply multiplying its previous probability with the probability of the new or updated child. In case of removing a context, the probability is divided by the probability of the new or updated child. Because the recalculation only takes place in part of the tree and allows us to preserver previous calculations for later use, it is partial. And because this recalculation only happens in trees that contain the added or removed context as a leaf, the calculation is incremental. We illustrate the PPC process using an extension of the hidden Markov model from the case study (See Fig.~\ref{fig:fuzzy2}). 
\begin{figure}[htb]
  \centering
  \includegraphics[width=3.5in]{fuzzy2}
  \caption{Fuzzy ontology specifying the Lecturing activity and the probabilities of different contexts in the presence of this higher-level context.}
  \label{fig:fuzzy2}
\end{figure}
In this figure the usesSlides context is added to the observation set at some point. This change in the context set requires us to recalculate the probabilities of the higher-level contexts. For this we use PPC: we recalculate only the values of the ancestors if the usesSlides context (c.q. isPresenting and isLecturing). The probability of isPresenting is $.85 * .90$, because the usesSlides context is its only child. The probability of isLecturing has to be recalculated. This can be done by simply multiplying its previous probability with the probability of the new child isPresenting: $isLecturing = 0.0004104 * .90 * .85 = 0.000313956$\footnote{Note that this value has to be multiplied with a constant in order to get the real probability value}.

@@

\section{Discussion}
In this article we started with presenting partial constraint checking: an efficient algorithm for timely detecting inconsistencies in huge sets of contexts which change frequently. Partial constraint checking was 15 times more efficient than the traditional greedy approach and reduced the conflict detection miss rate from 52.2\% to 0.1\%. This algorithm is clearly a dramatic improvement of the traditional approach and vital to any context aware application that has to handle frequently generated contexts.

>>>>

%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
\keywords{Pervasive Computing, Context Ontology Model, Partial Constraint Checking, Context Lifecycle, Context Inconsistency Resolution, Fuzzy OWL.}

@@

In the next chapter we start with an overview of the existing solutions for handling context inconsistencies and argue why these solutions do not meet the requirements of pervasive applications. Next we present Partial Constraint Checking: an algorithm for constraint checking on huge sets of contexts that is far more efficient and effective than the existing approaches. In addition we discuss a Context Inconsistency Resolution algorithm that allows for resolving conflicting context sets without human intervention. Finally we propose a whole new approach to context inconsistencies: incorporating inconsistencies in to the context model by using fuzzy ontologies. We show that this approach makes conflict detection and resolution superfluous and has a lot of potential for an efficient implementation using Partial Probability Calculation.

@@

Context management for consistency however has not been adequately studied in the existing literature. None of the studies on context-awareness discusses a way for detecting context inconsistencies for reliable pervasive computing~\cite{xu:2010:PCC, bu:2006:CCM}. Even though other disciplines as artificial intelligence and software engineering conducted related research, it does not provide adequate support for context inconsistency detection in ubiquitous computing. In addition the strategies proposed in literature for resolving context conflicts are not suited for pervasive computing. Some are based on assumptions that may not apply to general pervasive environments. Other require human participation for conflict resolution, which is usually expensive and slow for pervasive computing~\cite{xu:2010:PCC}. Finally no research has been done on the potential of fuzzy ontologies to represent inconsistencies in the environment or the perception of this environment instead of trying to resolve them~\cite{ko:2009:IOFO}.

@@

\section{Partial Constraint Checking}
Constraint checking techniques have been extensively studied in software engineering. Existing constraint checking algorithms focus on checking software artifacts that do not change rapidly or frequently~\cite{xu:2010:PCC}. Context-aware applications require more efficient algorithms because they use a huge set of contexts, which can change very rapidly and frequently (in the range of milliseconds). An inefficient software solution for this does not only require massive computing power, but interestingly also induces a higher inconsistency detection miss rate. Because of the computing delay conflicting contexts slip through the context buffer before they are detected by the software~\cite{xu:2010:PCC}. One example of an inefficient approach is non-incremental checking: whenever there is a change in the set of software artifacts these artifacts are each checked against the entire set of consistency constraints to find out all detectable inconsistencies. An improvement to this would be incremental checking: only a subset of all constraints are checked, namely those that are affected by the specific change in the artifact set. But Xu et \textit{al.} made the most substantial progress by replacing the traditional entire constraint checking approach by partial constraint checking based on a consistency computation tree~\cite{xu:2010:PCC}. Their idea is that constraints can be represented as trees with nodes for logical operations and leaves for specific properties of contexts or context sets. Whenever a context is added to or deleted from a context set in time, the branch corresponding to this context can respectively be added to or deleted from the tree (see Fig.~\ref{fig:cct}). 
\begin{figure}[htb]
  \centering
  \includegraphics[width=3.5in]{cons_comp_tree}
  \caption{Consistency Computation Tree: branches can be added or deleted to represent changes in a context set. Source:~\cite{xu:2010:PCC}}
  \label{fig:cct}
\end{figure}
Because intermediate values are retained in the tree nodes after calculation, they can be reused whenever the tree changes. More specifically, when a branch is added, only the values of the new branch itself and of the nodes from the branch top to the tree root need to be calculated. When a branch is deleted, only the values for the nodes from the branch top to the tree root have to be recalculated.

@@

\section{Context Inconsistency Resolution}
Now that we have an efficient method for detecting inconsistencies, the task of resolving the conflicts remains. As discussed above the strategies for conflict resolution in literature are not well suited for pervasive computing because their assumptions do not apply or they require human participation~\cite{xu:2010:PCC}. What is really needed is an algorithm that in the case of a conflict between two or more contexts decides which context has the highest reliability and retains that context. Bu et \textit{al.} show us that this is possible through extending the ontology based context model with additional information about the context's status and temporal properties~\cite{bu:2006:CCM}. More specifically they propose an algorithm that retains for every conflict set the context with the highest \textit{relative frequency}: frequency of context updates relative to update interval and context age (see Fig.~\ref{fig:cir}). 
\begin{figure}[htb]
  \centering
  \includegraphics[width=3.5in]{cir}
  \caption{Context inconsistency resolution in action. Source:~\cite{bu:2006:CCM}}
  \label{fig:cir}
\end{figure}
This is based on the assumption that contexts that are most stably perceived by a system are most likely to be correct. This assumption is applicable to most of the environments in which pervasive systems run and is applied in many domains as a domain specific approach~\cite{xu:2010:PCC}.

@@

\subsection{Case Study}
Suppose we want to build a context-aware application that has a certain notion of teacher's activities to assist them in the best possible way (for instance automatically opening slides when a lecture starts). In the traditional approach we could build an ontology specifying where lectures take place and that a lecture starts when the teacher appears at the lectern. If now an unpredicted observation is made (sensor cross read or teacher momentarily leaves the room), an inconsistency occurs that has to be resolved for the application to proceed with its normal operation. In the fuzzy ontology approach these unpredictable events can be modeled so that the ontology is more robust and always yields the most apt interpretation of the environment without the need for conflict resolution. A simplified hidden Markov model for this ontology is depicted in Fig.~\ref{fig:fuzzy}.
\begin{figure}[htb]
  \centering
  \includegraphics[width=3.5in]{fuzzy}
  \caption{Fuzzy ontology specifying a teacher's activities and the probabilities of different locations in the presence of these higher-level context.}
  \label{fig:fuzzy}
\end{figure}
As the numbers in this model indicate, the most probable location of a teacher while lecturing is at the lectern (L). In the aisle (A) is less probable and at the office (O) is very unlikely. For the activity of performing other work (denoted as Working in the model), the probabilities are the other way around. This fuzzy ontology model can now be used to calculate the relative probability of each higher-level context (c.q. Lecturing and Working) based on a set of low level contexts observed within a certain time window. A possible set of observations could be [$L_1$, $L_2$, $L_3$, $A_1$, $A_2$]: the teacher was observed three times at the lectern and twice in the aisle in this time window. The probability of the Lecturing activity can now be calculated by multiplying the probabilities of the observations in the presence of this higher level context\footnote{We multiply the results with a constant $c$ to obtain a valid probability value in the end.}: $pLecturing = .80 * .80 * .80 * .15 * .15 * c_1 = .01152 * c_1 = 0.9998$. The probability for Working can be calculated in a similar way: $pWorking = .05 * .05 * .05 * .15 * .15 * c_1 = 0.000002813 * c_1 = 0.0002$. The probabilities indicate that the teacher is still lecturing, even though the teacher being observed in the aisle conflicts with being observed at the lectern. Another example of observations could be [$L_1$, $A_1$, $A_2$, $O_1$, $O_2$]: the teacher is observed at the lectern once, twice in the aisle and twice at the office. The probabilities for Lecturing and Working are calculated as follows: $pLecturing = .80 * .15 * .15 * .05 * .05 * c_2 = 0.000045 * c_2 = 0.059$ and $pWorking = .05 * .15 * .15 * .80 * .80 * c_2 = 0.00072 * c_2 = 0.941$. The probabilities now indicate that the teacher is doing work other than lecturing, even though the observation of the teacher at the lectern is definitely conflicting with the observations at the office.
\subsection{Results}
The case study shows that a fuzzy ontology allows us to always define the most probable higher-level context given the observations, regardless of any inconsistencies in the observed context set and without the need for conflict detection and resolution.

Another observation we can do is that the hidden Markov model for a fuzzy ontology has a tree-like structure. This structure allows for very efficient incremental Partial Probability Calculation (PPC). The PCC principle works as follows: the probability of a high-level context is calculated by multiplying the probabilities of its children contexts. When a new context is added to or removed from the observation set and thus from the tree, only the probabilities of the observation's ancestors in the tree up to the root have to be recalculated. In case of adding a context, the probability of an ancestor in the tree can be recalculated by simply multiplying its previous probability with the probability of the new or updated child. In case of removing a context, the probability is divided by the probability of the new or updated child. Because the recalculation only takes place in part of the tree and allows us to preserve previous calculations for later use, it is partial. And because this recalculation only happens in trees that contain the added or removed context as a leaf, the calculation is incremental. We illustrate the PPC process using an extension of the hidden Markov model from the case study (See Fig.~\ref{fig:fuzzy2}). 
\begin{figure}[htb]
  \centering
  \includegraphics[width=3.5in]{fuzzy2}
  \caption{Fuzzy ontology specifying the Lecturing activity and the probabilities of different contexts in the presence of this higher-level context.}
  \label{fig:fuzzy2}
\end{figure}
In this figure the usesSlides context is added to the observation set at some point. This change in the context set requires us to recalculate the probabilities of the higher-level contexts. For this we use PPC: we recalculate only the values of the ancestors if the usesSlides context (c.q. isPresenting and isLecturing). The probability of isPresenting is $.85 * .90$, because the usesSlides context is its only child. The probability of isLecturing has to be recalculated. This can be done by simply multiplying its previous probability with the probability of the new child isPresenting: $isLecturing = 0.0004104 * .90 * .85 = 0.000313956$\footnote{Note that this value has to be multiplied with a constant in order to get the real probability value}.

@@

\section{Discussion}
In this article we started with presenting partial constraint checking: an efficient algorithm for timely detecting inconsistencies in huge sets of contexts which change frequently. Partial constraint checking was 15 times more efficient than the traditional greedy approach and reduced the conflict detection miss rate from 52.2\% to 0.1\%. This algorithm is clearly a dramatic improvement of the traditional approach and vital to any context-aware application that has to handle frequently generated contexts.
